\textbf{Optimization of Porous Silicon Gradient Refractive Index Structures.}

### 2020-06-13 paper.tex
Review manuscript. Correct mistakes. Start on experimental details

\begin{abstract}
We propose a simple setup which allows the calculation of the current
density as a function of position within an electrochemical cell used
to produce porous Silicon structures with a controlled gradient in their
index of refraction. We produced samples under several conditions and
characterized them with scanning electron microscopy, and
ultraviolet-visible-near infrared reflectance spectroscopy. Our
results allow us to obtain from experiments on a single or just a few
samples a full characterization of the dependence of porosity,
dielectric function and speed of attack on the current density and on
position, so that novel nonhomogeneous devices may be designed and
built.

@@

\section{Theory}
\subsection{Calculation of the current density}
\notaL{check where acronyms are first defined}
A gradient in the optical properties of porous Silicon (PS) systems may be
obtained by electrochemically attacking a Si surface with a position dependent current
density ${\bm j(\bm r)}$, as the porosity and refractive index depend on
the current density, as well as the speed of attack and the width of the
porous layer. Nevertheless, it is not a simple to task to measure the
{\em local} value of $\bm j(\bm r)$ along the surface. This hampers the
design and optimization of well characterized reproducible GRIN PS
structures. Thus, a system for which the current density field may be
easily computed would be useful.
Though $\bm j(\bm r)$ may be in principle be numerically calculated for a
given arbitrary experimental setup,  here we propose a particularly
simple setup that allows an expression for the current in terms
of a series that is rapidly convergent and each of whose terms is a
simple analytic function. With this setup it is easy to characterize
the dependence of different properties of interest on the current, and
this allows the design and optimization of diverse GRIN PS devices.

We assume that our electrolytic cell has the shape of a
prism (see Fig. \ref{f:celda}) with a horizontal rectangular base of sides $a$ and $b$, and
filled with the electrolyte up to a height $c$. We assume that the
walls of the cell are insulating while the bottom is completely
covered with the sample, which is a good conductor which is
electrically grounded. Within the electrolyte we position an electrode
which we assume is a thin conductor, insulated from the electrolyte
but for a small area near its tip, which we thus approximate as a
point current source.
\begin{figure}
  \centering
  \includegraphics[scale=.4]{Images/celdan}
  \caption{Prism shaped electrolytic cell with a rectangular base of
    sides $a$, $b$ filled with electrolyte up to a height $c$. The
    walls of the cell are insulating and the bottom is completely
    covered by the sample, which is assumed to be a good
    conductor. The tip of the electrode is a thin insulated wire with
    just the tip uncovered and located at $ \vec{r}_0 $ within the
    liquid.\notaL{Quité el tipo de letra del pie de figura. Que la pongan en la
      revista.}\notaL{En la punta pusiste dos esferas. Quítalas y sólo
    pon un puntito dorado del ancho del cable. Muestra que la base
    está aterrizada}\notaL{Indica el voltaje y la
    corriente}\notaL{¿Está bien la polaridad de la batería?}}
  \label{f:celda}
\end{figure}
Within the electrolyte the current density is
\begin{equation}
  \label{eq:j}
  \bm j=\sigma\bm E,
\end{equation}
with
$\sigma$ the conductivity, $\bm E=-\nabla\varphi$ the electric field,
and $\varphi$ the electric potential. For a
stationary situation\notaL{Latex usa nabla como el nombre de
  $\nabla$}
$\nabla\cdot \bm j = \sigma\nabla\cdot\bm E=0$. Thus, within the electrolyte,
\begin{equation}
  \label{eq:laplace}
\nabla^{2} \varphi=0.
\end{equation}
Consider now an arbitrary closed surface
$\mathcal S$
within the electrolyte that surrounds completely the tip of the
electrode. Integrating $ \bm j$ over this surface we obtain
\begin{equation}
  \label{eq:I}
\int_{\mathcal S'} d\bm a \cdot \bm j = -I,
\end{equation}
where the prime on $\mathcal S'$ means we remove from the surface a
very small hole through which the wire that feeds the current gets
through. Assuming the wire is narrower than any other relevant
distance in the system, we may interpret Eq. \eqref{eq:I} as an
integral over a closed surface of the current \eqref{eq:j} within the
electrolyte, ignoring the actual current within the wire. Thus
\begin{equation}
  \label{eq:gauss}
  \int_{\mathcal S} d\bm a \cdot \bm E = -\frac{I}{\sigma}
\end{equation}
Using Gauss's law, we interpret this equation as a source for the
potential in the form of a point charge
\begin{equation}
  q_{_{0}} = -\frac{I}{4 \pi \sigma}
\end{equation}
at the position $\vec{r}_{_{0}}=(x_{_{0}}, y_{_{0}}, z_{_{0}})$ of the
tip of the electrode. Note that $q_0$ would be the total charge,
and it shouldn't be further screened through the permittivity of the
electrolyte.

@@

\notaL{ref}. The potential within the electrolytic cell coincides the
the field within an infinite vacuum containing the point charge

@@

    charges at $(\pm x_0, \pm y_0, z_0)$.}

@@

four equal charges situated at the positions $\bm r_0$, $\bm

@@

The electrostatic potential produced by an infinite array of
charges expressed as a sum in real space of Coulomb terms has ill
convergence properties. The potential can be obtained
using the Ewald summation technique \notaL{ref}, which yields two rapidly converging
series, one in real space and another one in reciprocal space.
Nevertheless, as we need the potential and the field to obtain the
current density on the surface of the sample on which there are no
real nor image charges, we may perform a somewhat simpler planewise
summation, where we first obtain the potential and field produced by a
simple periodic lattice of charges using a fast convergent sum over the 2D
reciprocal vectors, and then sum over the charges that form the basis
of the 2D {\em crystal} plane and finally sum over all the planes \notaL{ref}.

We consider first a system made of identical unit charges
ocuppying the positions $\bm R$ of a 2D Bravais lattice $\{\bm R\}$,
which we will take as a rectangular lattice with lattice parameters
$2a$ and $2b$, and lying, for the time being, at the $z=0$ plane.
The potential it produces obeys
\begin{equation}
  \label{eq:poissonRed}
  \nabla^2\varphi(\bm r)=-4\pi\sum_{\bm R}\delta(\bm r_\|-\bm
  R)\delta(z)=-\frac{4\pi}{A}\sum_{\bm G}e^{i\bm G\cdot\bm r_\|}\delta(z),
\end{equation}
where $\bm r_\|$ is the projection of the observation position $\bm
r=(x,y,z)$ onto the $xy$ plane,
$\delta(\ldots)$ is Dirac's delta function, $\{\bm G\}$ is the
reciprocal lattice defined through $e^{i\bm G\cdot\bm R}=1$ for all
$\bm R$ and $\bm G$, $A=4ab$ is the area of the unit cell
and we used the Fourier representation of the 2D
periodically repeated delta function. We introduce a 2D Fourier representation for
the potential \notaL{Checar consistencia entre phi y varphi en todos lados}
\begin{equation}
  \label{eq:phifourier}
  \phi(\bm r)=\sum_{\bm G}\phi_{\bm G}(z)e^{i\bm G\cdot\bm r_\|}
\end{equation}
where $\phi_{\bm G}(z)$ is the Fourier coefficient of the potential at
the height $z$. Substitution in Eq. \eqref{eq:poissonRed} yields the
ordinary differential equation \notaL{El signo estaba mal. Chécalo.}
 \begin{eqnarray}
 \frac{d^2}{dz^2}\phi_{\bm G}(z)-G^2\phi_{_{G}}(z)=
   -\frac{4 \pi}{A} \delta(z),
 \end{eqnarray}
an homogeneous equation for $z\ne0$ which may be trivially
solved. After applying boundary conditions at $z=0$ and regularity
conditions at infinity, we obtain
\begin{equation}
  \label{eq:phiG}
  \phi_{\bm G}(z)=
  \begin{cases}
    -\dfrac{2\pi}{A}\abs{z}&\text{if }G=0,\\
    \dfrac{2 \pi}{AG}  e^{- G \abs{z}}&\text{if }G\neq 0.
  \end{cases}
\end{equation}
The case $G=0$ corresponds to the potential produced by a uniformly charged plane,
while the case $G\ne 0$ is the symmetric solution that decays far from
the source plane. Substituting into Eq. \eqref{eq:phifourier} yields
\begin{eqnarray}
  \label{eq:phiG1}
 \phi(\bm r) = -\frac{2 \pi}{A}  \abs{z} +{\sum_{\bm G}}' \frac{2 \pi
   }{AG} e^{- G \abs{z}} e^{i \bm G\cdot\bm r_\|},
 \end{eqnarray}
where the prime indicates that the term $G=0$ should be ommited from
the sum.

We consider now the basis of our 2D lattice, with equal charges at the
four positions $(\pm x_0, \pm y_0)$. Each of the corresponding four
sublattices produces a potential as that in Eq. \eqref{eq:phiG1} but
shifted by the corresponding basis vector, yielding
\begin{equation}
  \label{eq:phiG4}
  \phi(\bm r) = -\frac{8\pi}{A} \abs{z}
    +{\sum_{\bm G}}' \frac{8\pi}{AG} e^{- G \abs{z}}\cos(G_x x_0)
    \cos(G_y y_0) e^{i \bm G\cdot\bm r_\|}.
\end{equation}
Finally, we shift the origin of Eq. \eqref{eq:phiG4} vertically to each of the planes shown in
Fig. \ref{fig:sideview} and multiply by the corresponding charge $q_0$
or $-q_0$ to obtain the contribution to the total potential. The charge $q_0$ corresponds to
the planes with heights of the form $z_0+4nc$ and $-z_0+(4n+2)c$, with
$n=\ldots-2,-1,0,1,2\ldots$ an arbitrary integer, while the charge
$-q_0$ corresponds to the planes with height $z_0+(4n+2)c$ and
$-z_0+(4n+4)c$.

We concentrate our attention only on the region $a\le z<z_0$. From
Fig. \ref{fig:sideview} we see that the $G=0$
field has contributions only from the charges at $z_0$ and their
images ant $-z_0$, this contribution is like that of a parallel plate
capacitor $E_{0z}=-4\pi q_0/ab=-16\pi q_0/A$, which corresponds to the
potential $\phi_0=16\pi q_0 z/A$. Other contiguous planes of with opposite charges
contribute to the field elsewhere. The contributions to the potential
from the $G\ne0$ are simple geometric sums that may be summed
analytically over $n$ and yield the total potential
\begin{equation}
  \label{eq:phitot}
  \begin{split}
    \phi(\bm r)=&\frac{16\pi q_0}{A}\biggl(z+{\sum_{\bm G}}'\frac{2\sinh
      Gc}{G \sinh 2Gc}\cosh G(c-z_0)\cos G_x x_0 \cos G_y y_0\\
    &\times \sinh Gz\, e^{i\bm G\cdot\bm r_\|}\biggr).
  \end{split}
\end{equation}

Notice that the terms of the sum above converge exponentially to zero
as $G$ increases for any $z$ such that $0\le z <z_0 \le c$, and thus
the sum is convergent. The divergence in the limit $ z
\longrightarrow z_0 $ is not worrisome, as we are interested in the
field close to bottom $z=0$ of the cell.

The current density normal to the surface may now be obtained as
$j_\perp=-\sigma\partial\phi/\partial z$,
\begin{equation}
  \label{Eq:J}
  \begin{split}
    j_\perp(\bm r) =& -\frac{I}{ab}\biggl(1+{\sum_{\bm G}}'\frac{2\sinh
      Gc}{\sinh 2Gc}\cosh G(c-z_0)\cos G_x x_0 \cos G_y y_0\\
    &\times \cosh Gz\, e^{i\bm G\cdot\bm r_\|}\biggr).
  \end{split}
\end{equation}
Eq. \eqref{Eq:J} is a very rapidly convergent expression that allows,
in particular, to calculate the current $j_\perp(x,y,0)$ that attacks
the substrate to produce the PS sample.
Integrating over the surface of the sample we verify
\begin{equation}
  \label{eq:intj}
  \int_0^adx\int_0^bdy\,j_\perp(\bm r)=-I,
\end{equation}
as the contributions of opposite non-null reciprocal vectors cancel
out. As expected, all of the current that leaves the electrode find
its way to the substrate.
\notaL{En este sistema, ¿cuál es el cátodo y cuál el ánodo? ¿de dónde
  hacia dónde va la corriente? ¿$q_0$ es positiva o negativa?}

@@

\caption{\notaL{reordené figura y rehice} Experimental set-up for
      the fabrication of porous silicon with GRIN, consisting of a
      rectangular electrochemical cell. We indicate the Si wafer (dark
      gray), the teflon cell (light gray), the sealing o-ring (black
      circles), the insulated feeding cable (white lines within black
      lines), the uncovered platinum tip (white) within the
      electrolyte (blue), the electrical contact below the sample
      (yellow) and the current source. We indicate some geometrical
      parameters of the arrangement $x0$, $z0$, $a$ and $c$
      and the current source (upper panel) and a few current density
      field lines. Top (middle panel) and
      side (lower panel) schematic views of the resulting sample,
      illustrating three regions $A$, $B$ and $C$ of relatively high, medium and
      low porosity respectively illustrating the gradients in pore
      size, density and depth. \notaL{Es $\nabla p$ razonable para el
        gradiente de la porosidad}}
    \label{fig:DE1}
  \end{figure}

In Fig. \ref{fig:DE1} we show a schematic representation of our
experimental setup to manufacture porous silicon with GRIN.
\notaL{Cambié las figuras} \notaL{La figura está feita. Hay un abultamiento extraño en la
  punta. Debería verse el aislamiento y la punta descubierta y más
  sumergida. La letra está demasiado chica. El gradiente de porosidad va
  en la dirección opuesta. El panel (a) tiene una discontinuidad. ¿Tú
  la hiciste o la adaptaste? Usa dos paréntesis (a), (b),
  (c). Alínealos. Pon primero la preparación y luego los resultados
  vistos de arriba y de lado. Está en desorden. Se ve muy descuidada. No dice nada sobre la
  preparación que no esté en la figura 1. Dos páneles no ilustran el
  método experimental, sino el resultado. Quizás yo la quitaría.}
As dicussed above, we fabricated an electrolytic cell with the shape of a
rectangular prism on the bottom of which we place a Si which makes an
electrical contact to a gold plate \notaL{Lo inventé; de qué material es
  el contacto bajo la muestra} which together make the anode. The
catode consists of a platinum wire insulated except for a small region
at its tip, which we take as a point current drain.
We also show a schematic drawing of the expected GRIN PS structures,
with a gradient in the porosity, and in the width, density and depth
of the resulting pores.

\subsection{Fabrication of Porous Silicon Monolayers GRIN}
Some of the simulated photonic structures were made by electrochemical
anodization on an oriented (100) p-type crystalline Si wafer
(resistivity 0.002 - 0.005 $ \Omega \cdot cm $), under galvanostatic
conditions. The electrochemical anodization process was performed at
room temperature, with an electrolytic mixture of aqueous hydrofluoric
acid (HF) (concentration: 48 $ \% $ wt), glycerol (purity: 99.8 $ \% $
wt) and ethanol (purity: 99.9 $ \%$ wt ) in 3: 7: 1 volume ratio,
respectively. After the anodizing process, the samples are rinsed with
ethanol (purity: 99.9 $ \% $  wt). The layout of the cell used is as
seen in Fig. \ref{fig:DE1} b), with an area of $ 1,206 \ \ cm^2 $
(this is the area attacked in the process), has a rectangular shape ($
0.62 \ \ cm \times 2.03 \ \ cm $)  so that the current densities  that
attacks has a lateral gradient with a greater surface area. They were
manufactured with a fixed current of $ 5 \ \ mA $ to generate
different current densities ranging from $ 6  $ to $ 2.5 \ \ mA / cm^2
$. Similarly, we use the current from $ 40 \ \ mA $ to generate $ 50 $
to $ 20 \ \ mA / cm ^ 2 $. The cathode is made of a platinum wire that
was arranged in the form of a tip with an approximate diameter of ($
0.5 \ \ mm $).
\subsection{Reflectance Measurement}
In the experiemental parte the reflectivity measurements were carried
out with a Perkin Elmer Lambda 950 UV-Vis-NIR spectrophotometer. To
calculate the complex refractive index $ \tilde{n}_{_{PS}} = n_{_
  {PS}} - ik_{_{PS}} $ of a porous layer with homogeneous porosity
and thickness with smooth interfaces to throughout the entire
structure, the more suitable is the adjustment procedure of the
experimental reflectance or transmittance spectrum, this is for the
theoretical part. In the case of the reflectance spectrum. The
corresponding theoretical calculations were carried using procedure
for single  smooth layer (monolayer).
 \begin{eqnarray}\label{Eq:ECMR}
R=\dfrac{r _ {_ {01}}r _ {_ {12}} e^{-2i\delta}}{1+r _ {_ {01}}r _ {_ {12}} e^{-2i\delta}}
\end{eqnarray}
where $ r_{_ {01}} $ y $ r_{_ {12}} $ are the Fresnel reflectance
coefficients at the interfaces $aire/PS$ y $PS/c-Si$ which are
defined in terms of their complex refractive index ($\tilde{n}_{_ {1}}
$, $\tilde{n}_{_ {2}} $ y $ \tilde{n}_{_ {3}} $), for them
$r_{_{ij}}=(\tilde{n}_{_{i}}\cos\theta_i-\tilde{n}_{_{j}}\cos\theta_j)/(\tilde{n}_{_{i}}\cos\theta_i+\tilde{n}_{_{j}}\cos\theta_j)$,
($j=i=0,1,2$), whilst $\delta=2\pi
d/\lambda\sqrt{\tilde{n}^2_{_{2}}-\tilde{n}^2_{_{1}}\sin \theta_1}$.
 To adjust the reflectance spectrum, the layer thickness is an input
 parameter in the Eq.\ref{Eq:ECMR}. The advantage of this method is
 the possibility of taking into account the interface roughness effect
 by including the Davies-Bennett factor in the coefficient of each
 Fresnel \cite{I17, I18, I19}, which represents the roughness of a
 Gaussian distribution.
\subsection{Morphological studies}
The morphologies of the etched porous layers (cross-sectional) was
observed using SEM Hitachi SU1510 scanning electron microscopes
(Hitachi High Technologies Canada, Inc., Toronto, Canada), the
thickness  and porosity  of each fabricated  film was measured.

>>>>

\begin{abstract}
We propose a simple setup which that the calculation of the current
density as a function of position within an electrochemical cell used
to produce porous Silicon structures with a controlled gradient in their
index of refraction. We produced samples under several conditions and
characterized them with scanning electron microscopy, and
ultraviolet-visible-near infrared reflectance spectroscopy. Our
results allow us to obtain from experiments on a single or just a few
samples a full characterization of the dependence of porosity,
dielectric function and speed of attack on the current density and on
position, so that novel nonhomogeneous devices may be designed and
built.

@@

\section{Theory}
\subsection{Calculation of the current density}
\notaL{check where acronyms are first defined}
A gradient in the optical properties of porous Silicon (PS) systems may be
obtained by electrochemically attacking a Si surface with a position dependent current
density ${\bm j(\bm r)}$, as the porosity and refractive index depend on
the current density, as well as the speed of attack and the width of the
porous layer. Nevertheless, it is not a simple task to measure the
{\em local} value of $\bm j(\bm r)$ along the surface. This hampers the
design and optimization of well characterized reproducible GRIN PS
structures.\notaL{Tenemos referencias para afirmar esto?} Thus, we
believe a
system for which the current density field may be
easily computed would be useful.
Though $\bm j(\bm r)$ may be in principle be numerically calculated for a
given arbitrary experimental setup,  here we propose a particularly
simple setup that allows an expression for the current in terms
of a series that is rapidly convergent and each of whose terms is a
simple analytic function. With this setup it is easy to characterize
the dependence of different properties of interest on the current, and
this allows the design and optimization of diverse GRIN PS devices.

We assume that our electrolytic cell has the shape of a
prism (see Fig. \ref{f:celda}) with a horizontal rectangular base of sides $a$ and $b$, and
filled with the electrolyte up to a height $c$. We assume that the
walls of the cell are insulating while the bottom is completely
covered with the sample, which is a which is an
electrically grounded good conductor. Within the electrolyte we position an electrode
which we assume is a thin conductor, insulated from the electrolyte
but for a small area near its tip, which we thus approximate as a
point current source.
\begin{figure}
  \centering
  \includegraphics[scale=.4]{Images/celdan}
  \caption{Prism shaped electrolytic cell with a rectangular base of
    sides $a$, $b$ filled with electrolyte up to a height $c$. The
    walls of the cell are insulating and the bottom is completely
    covered by the sample, which is assumed to be a good
    conductor. The tip of the electrode is a thin insulated wire with
    just the tip uncovered and located at $\bm r_0=(x_0,y_0,z_0)$ within the
    liquid.\notaL{Quité el tipo de letra del pie de figura. Que la pongan en la
      revista.}\notaL{En la punta pusiste dos esferas. Quítalas y sólo
    pon un puntito dorado del ancho del cable. Muestra que la base
    está aterrizada}\notaL{Indica el voltaje y la
    corriente}\notaL{Mejor pon una fuente de corriente en lugar de una
  fuente de voltaje}}
  \label{f:celda}
\end{figure}
Within the electrolyte the current density is
\begin{equation}
  \label{eq:j}
  \bm j=\sigma\bm E,
\end{equation}
with
$\sigma$ the conductivity, $\bm E=-\nabla\varphi$ the electric field,
and $\varphi$ the electric potential. For a
stationary situation\notaL{Latex usa nabla como el nombre de
  $\nabla$}
$\nabla\cdot \bm j = \sigma\nabla\cdot\bm E=0$. Thus, within the electrolyte,
\begin{equation}
  \label{eq:laplace}
\nabla^{2} \varphi=0.
\end{equation}
Consider now an arbitrary closed surface
$\mathcal S$
within the electrolyte that surrounds completely the tip of the
electrode. Integrating $ \bm j$ over this surface we obtain
\begin{equation}
  \label{eq:I}
\int_{\mathcal S'} d\bm a \cdot \bm j = -I,
\end{equation}
where the prime on $\mathcal S'$ means we remove from the surface a
very small hole through which the wire that feeds the current to the
electrode gets
through. Assuming the wire is narrower than any other relevant
distance in the system, we may interpret Eq. \eqref{eq:I} as an
integral over a closed surface of the current \eqref{eq:j} within the
electrolyte, ignoring the actual current within the wire. Thus
\begin{equation}
  \label{eq:gauss}
  \int_{\mathcal S} d\bm a \cdot \bm E = -\frac{I}{\sigma}
\end{equation}
Using Gauss's law, we interpret this equation as a source for the
potential in the form of a point charge
\begin{equation}
  q_{_{0}} = -\frac{I}{4 \pi \sigma}
\end{equation}
at the position $\vec{r}_{_{0}}=(x_{_{0}}, y_{_{0}}, z_{_{0}})$ of the
tip of the electrode. Note that $q_0$ would be the total charge,
and it shouldn't be further screened through the permittivity of the
electrolyte.

@@

\notaL{ref}. The potential within the electrolytic cell coincides
within the region $0\le x\le a$, $0\le y \le b$, $0\le z\le c$
with that an infinite fictitious empty space but for the point charge

@@

    charges $q_0$ at $(\pm x_0, \pm y_0, z_0)$.}

@@

four equal charges $q_0$ situated at the positions $\bm r_0$, $\bm

@@

The electrostatic potential produced by an infinite array of
charges expressed as a sum in real space of Coulomb terms has ill
convergence properties. The potential can be obtained
using the Ewald summation technique \notaL{ref}, which yields two rapidly converging
series, one in real space and another one in reciprocal space.
Nevertheless, as we have to evaluate the potential and the field on
the bottom of the cell to obtain the current density which attacks the
surface of the sample, and as there are no
real nor image charges there, we may perform a somewhat simpler planewise
summation, where we first obtain the potential and field produced by a
simple periodic lattice of charges using a fast convergent sum over the 2D
reciprocal vectors, we then sum over the charges that form the basis
of the 2D {\em crystal} plane and we finally sum over all the planes \notaL{ref}.

We consider first a system made of identical unit charges
ocuppying the positions $\bm R$ of a 2D Bravais lattice $\{\bm R\}$,
which we will take as a rectangular lattice with lattice parameters
$2a$ and $2b$, and lying, for the time being, at the $z=0$ plane.
The potential it produces obeys
\begin{equation}
  \label{eq:poissonRed}
  \nabla^2\varphi(\bm r)=-4\pi\sum_{\bm R}\delta(\bm r_\|-\bm
  R)\delta(z)=-\frac{4\pi}{A}\sum_{\bm G}e^{i\bm G\cdot\bm r_\|}\delta(z),
\end{equation}
where $\bm r_\|$ is the projection of the observation position $\bm
r=(x,y,z)$ onto the $xy$ plane,
$\delta(\ldots)$ is Dirac's delta function, $\{\bm G\}$ is the
reciprocal lattice defined through $e^{i\bm G\cdot\bm R}=1$ for all
$\bm R$ and $\bm G$, $A=4ab$ is the area of the unit cell
and we used the Fourier representation of the 2D
periodically repeated delta function. We introduce a 2D Fourier representation for
the potential \notaL{Checar consistencia entre phi y varphi en todos lados}
\begin{equation}
  \label{eq:phifourier}
  \phi(\bm r)=\sum_{\bm G}\phi_{\bm G}(z)e^{i\bm G\cdot\bm r_\|}
\end{equation}
where $\phi_{\bm G}(z)$ is the Fourier coefficient of the potential at
the height $z$. Substitution in Eq. \eqref{eq:poissonRed} yields the
ordinary differential equation \notaL{El signo estaba mal. Chécalo.}
 \begin{eqnarray}
 \frac{d^2}{dz^2}\phi_{\bm G}(z)-G^2\phi_{_{G}}(z)=
   -\frac{4 \pi}{A} \delta(z),
 \end{eqnarray}
an homogeneous equation for $z\ne0$ which may be trivially
solved. After applying boundary conditions at $z=0$ and regularity
conditions at infinity, we obtain
\begin{equation}
  \label{eq:phiG}
  \phi_{\bm G}(z)=
  \begin{cases}
    -\dfrac{2\pi}{A}\abs{z}&\text{if }G=0,\\
    \dfrac{2 \pi}{AG}  e^{- G \abs{z}}&\text{if }G\neq 0.
  \end{cases}
\end{equation}
The case $G=0$ corresponds to the potential produced by a uniformly charged plane,
while the case $G\ne 0$ is the potential produced by a sinusoidal
charge density on a plane given by the
symmetric solution that decays exponentially as we get away from
the source plane. Substituting into Eq. \eqref{eq:phifourier} yields
\begin{eqnarray}
  \label{eq:phiG1}
 \phi(\bm r) = -\frac{2 \pi}{A}  \abs{z} +{\sum_{\bm G}}' \frac{2 \pi
   }{AG} e^{- G \abs{z}} e^{i \bm G\cdot\bm r_\|},
 \end{eqnarray}
where the prime indicates that the term $G=0$ should be ommited from
the sum.

We consider now the basis of our 2D lattice, with equal charges at the
four positions $(\pm x_0, \pm y_0,0)$. Each of the corresponding four
sublattices produces a potential as that in Eq. \eqref{eq:phiG1} but
shifted by the corresponding basis vector, yielding
\begin{equation}
  \label{eq:phiG4}
  \phi(\bm r) = -\frac{8\pi}{A} \abs{z}
    +{\sum_{\bm G}}' \frac{8\pi}{AG} e^{- G \abs{z}}\cos(G_x x_0)
    \cos(G_y y_0) e^{i \bm G\cdot\bm r_\|}.
\end{equation}
Finally, we shift the origin of Eq. \eqref{eq:phiG4} vertically to each of the planes shown in
Fig. \ref{fig:sideview} and multiply by the corresponding charge $q_0$
or $-q_0$ to obtain the contribution to the total potential. The charge $q_0$ corresponds to
the planes with heights of the form $z_0+4nc$ and $-z_0+(4n+2)c$, with
$n=\ldots-2,-1,0,1,2\ldots$ an arbitrary integer, while the charge
$-q_0$ corresponds to the planes with height $z_0+(4n+2)c$ and
$-z_0+(4n+4)c$.

We concentrate our attention only on the region $a\le z<z_0$. From
Fig. \ref{fig:sideview} we see that the $G=0$
field has contributions only from the charges at $z_0$ and their
images at $-z_0$, and that this contribution is like that of a parallel plate
capacitor $E_{0z}=-4\pi q_0/ab=-16\pi q_0/A$, which corresponds to the
potential $\phi_0=16\pi q_0 z/A$. Other contiguous planes of with opposite charges
contribute to the field elsewhere. The contributions to the potential
from the $G\ne0$ are simple geometric sums that may be added
analytically over $n$ and yield the total potential
\begin{equation}
  \label{eq:phitot}
  \begin{split}
    \phi(\bm r)=&\frac{16\pi q_0}{A}\biggl(z+{\sum_{\bm G}}'\frac{2\sinh
      Gc}{G \sinh 2Gc}\cosh G(c-z_0)\cos G_x x_0 \cos G_y y_0\\
    &\times \sinh Gz\, e^{i\bm G\cdot\bm r_\|}\biggr).
  \end{split}
\end{equation}

Notice that the terms of the sum above converge exponentially to zero
as $G$ increases for any $z$ such that $0\le z <z_0 \le c$, and thus
the sum is convergent. The divergence in the limit $ z
\longrightarrow z_0 $ corresponds to the singularity of the potential
at the position of a point charge, and it is not worrisome, as we are
interested in the field close to bottom $z=0$ of the cell.

The current density normal to the surface may now be obtained as
$j_\perp=-\sigma\partial\phi/\partial z$,
\begin{equation}
  \label{Eq:J}
  \begin{split}
    j_\perp(\bm r) =& \frac{I}{ab}\biggl(1+{\sum_{\bm G}}'\frac{2\sinh
      Gc}{\sinh 2Gc}\cosh G(c-z_0)\cos G_x x_0 \cos G_y y_0\\
    &\times \cosh Gz\, e^{i\bm G\cdot\bm r_\|}\biggr).
  \end{split}
\end{equation}
Eq. \eqref{Eq:J} is a very rapidly convergent expression that allows,
in particular, to calculate the current $j_\perp(x,y,0)$ that attacks
the substrate to produce the PS sample.
Integrating over the surface of the sample we verify
\begin{equation}
  \label{eq:intj}
  \int_0^adx\int_0^bdy\,j_\perp(\bm r)=I,
\end{equation}
as the contributions of opposite non-null reciprocal vectors cancel
out. As expected, all the current that leaves sample finds
its way to the cathode.

@@

\caption{\notaL{reordené figura y rehice} Experimental set-up for
      the fabrication of porous silicon with GRIN, consisting of a
      electrochemical cell with the shape of a
      rectangular prism. We indicate the Si wafer (dark
      gray), the teflon cell (light gray), the sealing o-ring (black
      circles), the insulated feeding cable (white lines within black
      lines), the uncovered platinum tip (white) within the
      electrolyte (blue), the electrical contact below the sample
      (yellow) and the controlled current source. We indicate some geometrical
      parameters of the arrangement $x_0$, $z_0$, $a$ and $c$
      and the current source (upper panel) and a few current density
      field lines. Top (middle panel) and
      side (lower panel) schematic views of the resulting sample,
      illustrating three regions $A$, $B$ and $C$ of relatively high, medium and
      low porosity respectively illustrating the gradients in pore
      size, density and depth. \notaL{Es $\nabla p$ razonable para el
        gradiente de la porosidad}}
    \label{fig:DE1}
  \end{figure}

In Fig. \ref{fig:DE1} we show a schematic representation of our
experimental setup to manufacture porous silicon with GRIN.
\notaL{Cambié las figuras}
As dicussed above, we fabricated an electrolytic cell with the shape of a
rectangular prism on the bottom of which we place a Si wafer which makes
electrical contact to a gold plate \notaL{Lo inventé; de qué material es
  el contacto bajo la muestra} which together make the anode. The
catode consists of a platinum wire insulated except for a small region
at its tip, which we take as a point current drain.
We also show a schematic drawing of the expected GRIN PS structures,
with a gradient in the porosity, and in the width, density and depth
of the resulting pores.

\subsection{Fabrication of PS  GRIN Monolayers}
We manufactured single layer PS structures through an electrochemical
anodization on a (100) $p$-type crystalline Si wafer
(resistivity 0.002 - 0.005 $ \Omega\,\text{cm}$), under galvanostatic
conditions. The process was performed at
room temperature,\notaL{Controlaste/mediste la temperatura?} with an
electrolytic mixture of aqueous hydrofluoric
acid (HF) (concentration: 48 $ \% $ wt), glycerol (purity: 99.8 $ \% $
wt) and ethanol (purity: 99.9 $ \%$ wt ) in a 3:7:1 volume ratio,
respectively. After the anodizing process, the samples were rinsed with
ethanol (purity: 99.9 $ \% $  wt). The electrolytic cell had the shape
of a rectangular prism as shown in Fig. \ref{fig:DE1}, with a base of
size $0.62\text{cm}\times2.03\text{cm}$) with an uncovered sample area
of $1.206\text{cm}^2$. \notaL{¿Cuáles fueron las coordenadas del
  electrodo?} We manufactured samples using a current $I=5\text{mA}$
which yielded current densities in the range
$j_\perp=2.5-6.0\text{mA}/\text{cm}^2$ and other smples using a larger
current $I=40\text{mA}$ corresponding to current densities in the
range $j_\perp=20-50\text{mA}/\text{cm}^2$. The cathode was a
platinum wire that with an approximate diameter of $0.5\text{mm}$.
\subsection{Reflectance Measurement}
We measured the reflectivity spectra using a {\em Perkin Elmer Lambda 950}
UV-Vis-NIR spectrophotometer.
